A community-maintained standard library of population genetic simulations

### 2019-12-18 popSimManu.tex
updates to two pop results

\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but specifying such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and the possibility for error.
Population genetics, as a field, also lacks standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, \stdpopsim, that attempts to rectify this situation.
\stdpopsim is a community-driven, open source project, which provides a standard
catalog of published simulation models from a wide range of organisms,
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature, community-based quality control pipeline.
We share some examples demonstrating how \stdpopsim
can be used to systematically compare demographic inference methods,
and encourange an even broader
community of developers to join in the effort to standardize
population genetic methods.}
\date{\today}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations 
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models}, 
the distribution of fitness effects \citep{kim2017inference},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247}, 
the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods which are validated by a set of bespoke simulations produced by individual
groups. Beyond a lack of efficiency from duplicated effort, such a situation leads
to decreases in reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby helping to usher in the next era of population genetic data analysis.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here

The degree to which modeling assumptions and different choices of summaries of the
data can affect the ultimate results has been challenging to assess. Yet, there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, \MSMC methods applied to human genomes have
% JK is this right? > 100 years doesn't sound very ancient to me.
suggested large ancient ($>100$ years ago) ancestral population sizes and
bottlenecks that have not been detected by SFS-based methods
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

@@

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community maintained catalog.
The \stdpopsim catalog currently contains five organisms: humans, orangutans,
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.

The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its accuracy.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate,
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give an
example of how it can be used to benchmark demographic inference methods.

@@

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution of the PopSim consortium currently is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{Drosophila},
\emph{Arabidopsis}, \emph{E. coli}, and \emph{Pongo}. A species definition consists
of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically-observed
heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of previously published maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic polymorphism.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig. \ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed very efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work, and authors
of simulation models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number, this citation information will result in a reasonably well-documented and reproducible simulation workflow.

\subsection*{The \stdpopsim Catalog}
The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}. These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components, and the three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture.
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture), and are therefore highly pertinent in the context of method development.

@@

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used 
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
competing methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairwayplot},
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the \cite{ragsdale2019models} model and using the
HapMapII genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under a sophisticated demographic model
and empirical genetic map. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC). The solid black lines represent the true effective population
size as inferred from the inverse coalescent rate defined by the model.
Blue lines represent estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
generally the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history---that inferred by \cite{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can even do away with the demographic model entirely, and simulate under a human
genome architecture and genetic map to see how well methods might do at recovering
the population history of a constant sized population. We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

@@

As most method development for population genetics has been focused on human-scale
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting, and generally worse than what we
observe under simulations of the human genome.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_popn_comp.png}
\caption{\textbf{Parameters estimated using a multi-population human model}.
Here we show estimates of $N(t)$ inferred using \dadi, \texttt{fastsimcoal2}, or \smcpp.
A) Data were generated by simulating
replicate human genomes under the \cite{gutenkunst2009inferring} model and using the genetic map
inferred in \cite{international2007second}. B) For \dadi and \texttt{fastsimcoal2} we show parameters inferred
by fitting the depicted IM model, which includes population sizes, migration rates, and a split
time between CEU and YRI samples. C) Population size estimates for each population (rows)
from \dadi, \texttt{fastsimcoal2}, and \smcpp (columns). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model. The population split date, $T_{div}$, is shown at
the bottom, with a common X-axis to the background figure.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\paragraph*{Multi-population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored parameter estimation of population divergence parameters. In particular
we performed simulations under multi-population models for humans and \emph{Drosophila}
and inferred simulated parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration. For human
models with more than two-populations, e.g. \cite{gutenkunst2009inferring},
this means that we are inferring parameters for a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B), however this is common practice in the field
when applying IM-type models.

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time, for each of the inference methods, using samples drawn from African and European populations
simulated under the \cite{gutenkunst2009inferring} model. Each of the methods compared has stronger
and weaker points to their accuracy across parameters. For instance the simple, IM-style
analyses we have performed do not capture recent population size change in CEU, however
they do a good job of estimating the simpler simulated YRI population size history from this model.

Again we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{Drosophila melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates for simulations drawn from
the model of \cite{li2006inferring}, which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population. These simulations are generated
using the recombination map of \cite{comeron2012many}. Here methods differ dramatically
in the estimates, with \smcpp yielding quite noisy estimates of population size
towards the more ancient past and all methods failing to capture the very brief
population bottleneck in Europe, although the IM model of course only allows for
two epochs in the population size history.

Again the goal of these analyses are not to perform thorough benchmarking of
the methods at hand under a wide range of simulations, but instead to show how
the tools that we have built could be used to facilitate such comparisons.

@@

The \stdpopsim library is the first piece of this effort, in that it allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
has yet to coalesce around a set of gold-standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology \citep{moult1995large} and machine learning \citep{russakovsky2015imagenet} have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and invaluable tradition in population genetics.

Moreover individual research groups are often in the position of having to independently
reimplement complex, previously published models. \stdpopsim offers solutions to both of
these issues by creating a quality controlled, easy-to-use, gateway to models and organisms
that are central to modern research in genetics.

We have built \stdpopsim, and the PopSim Consortium, to be an open, community-developed
project, meaning that it has excellent potential for future growth.
Thus far we have implemented genome representations and genetic maps for the three most
common study systems in computational genetics: humans, \emph{Drosophila}, and \emph{Arabidopsis}.
The standard operating procedures that we have outlined in our detailed documentation
means that new researchers, who are interested in adding their favorite organism/genome
are free to do so with very few barriers to entry (see Methods).
Further, new demographic models can be added
to each of the organisms in the catalog through similarly well documented steps.
A long-term goal then is that through a mechanism such as \stdpopsim, the community
can determine (in an open way) what is the correct implementation for a published demographic model,
and in turn, what are the models in which the community is most interested.

@@

\section*{Methods}
\subsection*{Generating the PopSim resource}

>>>>

\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal2}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

\abstract{The explosion in population genomic data demands ever more complex
modes of analysis, and increasingly these analyses depend on sophisticated simulations.
Recent advances in population genetic simulation have made it possible to
simulate large and complex models, but specifying such models for a particular
simulation engine remains a difficult and error-prone task.
Computational genetics researchers currently re-implement simulation models
independently, leading to duplication of effort and the possibility for error.
Population genetics, as a field, also lacks standard benchmarks by which new tools for inference might
be measured. Here we describe a new resource, \stdpopsim, that attempts to rectify this situation.
\stdpopsim is a community-driven open source project, which provides a standard
catalog of published simulation models from a wide range of organisms
and supports multiple simulation engine backends.
% PopSim assures correctness in our implementations through a mature, community-based quality control pipeline.
We share some examples demonstrating how \stdpopsim
can be used to systematically compare demographic inference methods,
and we encourage an even broader
community of developers to join in the effort to standardize
population genetic methods.}
\date{\today}

@@

\section*{Introduction}
While population genetics has always used statistical methods to make inferences from data,
the degree of sophistication of the questions, models, data, and computational approaches
used have all increased over the past decade. Currently there exist myriad computational methods
that can infer the histories of populations
\citep{gutenkunst2009inferring,li2011inference,excoffier2013robust,schiffels2014inferring, terhorst2017robust,ragsdale2019models},
the distribution of fitness effects \citep{kim2017inference},
recombination rates \citep{chan2012genome,lin2013fast,Adrion662247},
and the extent of positive selection in genome sequence data
\citep{eyre2009estimating, alachiotis2012omegaplus,degiorgio2016sweepfinder2,kern2018diplos,sugden2018localization}.
While en masse these methods have increased our understanding of the
impacts of genetic and evolutionary processes, very little has been done to systematically
benchmark the quality of inferences gleaned from computational population genetics.
As large databases of population genetic variation begin to be used to inform public health procedures,
the accuracy and quality of these inferences is becoming even more important.
This is in large part because ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement as to what should
constitute community standards or best practices. The general modus operandi to date has been to
introduce methods that are validated by a set of bespoke simulations produced by individual
groups. Beyond inefficiency from duplicated effort, such a situation leads
to decreased reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software
and select the best tool for a given task, thereby substantially increasing the reliability of population genetic inferences.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here
%RNG: Removed the sap. :-P Feel free to put it back if you like it better.

The degree to which modeling assumptions and choices of data summaries can affect the ultimate results of population genetic inference has been challenging to assess. Yet there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
% JK is this right? > 100 years doesn't sound very ancient to me.
% RNG: Fixed time and removed unexplained acronyms.
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

@@

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \texttt{stdpopsim}, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains five organisms: humans, orangutans,
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, the user can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.

The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its accuracy.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give an
example of how it can be used to benchmark demographic inference methods.

@@

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution of the PopSim consortium currently is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{D.~melanogaster}, \emph{A.~thaliana}, \emph{E.~coli}, and\emph{P.~abelii}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically-observed
heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of previously published maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic polymorphism.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig.~\ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and we have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work and authors
of simulation models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number, this citation information will result in a reasonably well-documented and reproducible simulation workflow.

\subsection*{The \stdpopsim Catalog}
The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}.
% RNG: Pongo added by submission?
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components, and the three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture.
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

@@

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
% RNG: Not competitors... We complement each other ;-)
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairwayplot}, and
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the \cite{ragsdale2019models} model and using the
HapMapII genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under an empirical genetic map and a model of human migration out of Africa that includes archaic admiture \citep{ragsdale2019models}. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC).
% RNG: Removed duplicated info from caption.
While there is variation in accuracy among methods, populations, and individual replicates,
generally the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history that does not include archaic admixture \citep{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

We can even do away with the demographic model entirely and simulate under a human
genome architecture and genetic map to see how well methods might do at recovering
the population history of a constant sized population. We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

@@

As most method development for population genetics has been focused on human
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{Drosophila melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting and generally worse than what we
observe under simulations of the human genome.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/homo_sapiens_two_popn_comp.pdf}
\caption{\textbf{Parameters estimated using a multi-population human model}.
Here we show estimates of $N(t)$ inferred using \dadi, \fastsimcoal, or \smcpp.
A) Data were generated by simulating
replicate human genomes under the \cite{gutenkunst2009inferring} model and using the genetic map
inferred in \cite{international2007second}. B) For \dadi and \fastsimcoal we show parameters inferred
by fitting the depicted IM model, which includes population sizes, migration rates, and a split
time between CEU and YRI samples. C) Population size estimates for each population (rows)
from \dadi, \texttt{fastsimcoal2}, and \smcpp (columns). In shades of blue we show the estimated
% RNG: Updated caption to reflect what is shown.
$N(t)$ trajectories for each replicate. In black we show the census population size from the simulated model.
The population split date, $T_{div}$, is shown at
the bottom, with a common X-axis to the population size panels.}
\label{fig:IM_popn_human}
\end{center}
\end{figure}

\paragraph*{Multi-population demographic models}
As \texttt{stdpopsim} implements multi-population demographic models, we also
explored two population inference with several popular methods. Specifically,
we performed simulations under multi-population models for humans and \emph{Drosophila}
and inferred demographic parameters using \dadi, \fastsimcoal, and \smcpp.
For simplicity, we conducted inference in \dadi and \fastsimcoal by fitting an isolation with migration (IM) model
with constant population sizes and bi-directional migration \citep{hey2004im}. Our motivation for using
an IM model was to mimic the approach often used on empirical datasets, where a relatively
simple model is fit that may not reflect the true underlying demography.
For human models with more than two-populations (e.g. \cite{gutenkunst2009inferring})
this means that we are inferring parameters for a model which itself does
not match the model from which the data were generated (Figure
\ref{fig:IM_popn_human}A and B).

In Figure \ref{fig:IM_popn_human}C we show estimates of population sizes and divergence
time, for each of the inference methods, using samples drawn from African and European populations
simulated under the \cite{gutenkunst2009inferring} model. Our results highlight many
of the strengths and weaknesses of the different types of methods we used.
For instance, the SFS-based approaches where we fit simple IM models do not capture
recent exponential growth in the CEU population, however do consistently recover the
simulated YRI population size history. Moreover, these approaches allow for estimating
migration rates (Figure \ref{fig:homsap_mig_rates}), also leading to more accurate inference
of divergence times. By contrast, \smcpp is much better at capturing the recent exponential
growth in the CEU population, though the inferred population sizes are in general more
noisy. In addition, the assumption of no migration by \smcpp leads to divergence time
estimates that are consistently underestimated.

Again, we can compare between species and look at the performance of these methods in
the context of a two-population model from \emph{Drosophila melanogaster}. Figure
\ref{fig:two_popn_fly} shows parameter estimates for simulations drawn from
the model of \cite{li2006inferring}, which includes
an ancestral population in Africa, then a population expansion with a population
split and bottleneck into a European population with no post-divergence migration.
Here again, we find that \dadi and \fastsimcoal are more consistent in their inferred
parameters, but ignore the brief population bottleneck in Europe. The inferred
demographic parameters from \smcpp are more noisy, though in some cases are better
able to capture the bottleneck in the Europe population.

Although these results do not represent an exhaustive benchmarking,
we have highlighted some of the strengths and weaknesses of these methods.
Future work should build on these results and undertake more in-depth comparisons
under a wider range of simulated demographic models.

@@

The \stdpopsim library is the first piece of this effort, in that it allows for rigorous
standardization of complex population genetic simulations. Population genetics, as a field,
% RNG: We should have more coalescence puns in the paper! :-P
has yet to coalesce around a set of gold-standards for the crucial task of method
evaluation, which in our field hinges on simulation. In contrast, other fields such as
structural biology \citep{moult1995large} and machine learning \citep{russakovsky2015imagenet} have a long track record
of standardized method testing. We hope that our efforts represent the beginning of what
will prove to be an equally longstanding and invaluable tradition in population genetics.

Moreover individual research groups are often in the position of having to independently
reimplement complex previously published models. \stdpopsim offers solutions to both of
these issues by creating a quality controlled, easy-to-use, gateway to models and organisms
that are central to modern research in genetics.

We have built \stdpopsim, and the PopSim Consortium, to be an open, community-developed
project, meaning that it has excellent potential for future growth.
Thus far we have implemented genome representations and genetic maps for the three most
common study systems in computational genetics: humans, \emph{Drosophila}, and \emph{Arabidopsis}.
% RNG: Needs to be updated for final count.
The standard operating procedures that we have outlined in our detailed documentation
mean that new researchers, who are interested in adding their favorite organism/genome
are free to do so with few barriers to entry (see Methods).
Further, new demographic models can be added
to each of the organisms in the catalog through similarly well documented steps.
A long-term goal is that the community
can determine (in an open way) what is the correct implementation for a published demographic model,
and in turn, what are the models in which the community is most interested.

@@

\section*{Methods}
%\subsection*{Generating the PopSim resource}
% RNG: No need to empty subsection.

\subsection*{Model QC procedures}
As a consortium we have agreed to a standardized procedure for model inclusion
into \stdpopsim that allows for rigorous quality control. Imagine Developer A
wants to introduce a new model into \stdpopsim. Developer A implements the
demographic model for the relevant organism along with clear documentation
of the model parameters and populations. This model is submitted as a Pull Request,
where it is evaluated by a reviewer and then included as `preliminary',
but is not linked to the online documentation nor the command line interface.
Inclusion of a preliminary model triggers a QC issue, and a second developer,
Developer B, then independently reimplements the model from the relevant
primary sources and adds an automatic unit test for equality between the
QC implementation and the preliminary production model. If the two
implementations are equal the original model is included in \stdpopsim.
If not, we move to an arbitration process whereby A and B first try
to work out the details of what went wrong. If that fails, the original
authors of the published model must be contacted
to resolve ambiguities. This process can lead to cycles of revision of
the implementation from Developers A and B but allows the project to assure
quality. Indeed we have already had a number of interesting cases of arbitration
to this point. Further details of our QC process can be found in our
\href{https://stdpopsim.readthedocs.io/en/latest/development.html#}{developer documentation}.

\subsection*{Workflow for analysis of simulated data}
To demonstrate the utility of \stdpopsim we created \texttt{Snakemake}
workflows \citep{koster2012snakemake} that perform demographic inference on
tree sequence output from our package using a few common software packages.
Our choice of \texttt{Snakemake} allows complete reproducibility of the
analyses shown, and the entire Analysis repository is available from
\url{https://github.com/popgensims/analysis}.
